Machine Learning Course Project
========================================================

##Load Data
Downloading the datasets from internet (source = http://groupware.les.inf.puc-rio.br/har)
and loading the Raw train and test datasets:
```{r cache=TRUE}
download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv',destfile='pml-training.csv',method='curl')
download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv',destfile='pml-testing.csv',method='curl')
trainingRAW <- read.csv('pml-training.csv', na.strings= c("NA",""))
testingRAW <- read.csv('pml-testing.csv', na.strings= c("NA",""))

```

##Cleaning and feature Selection
We discard from the datasets the columns with NAs and the next columns:
- timestamp and windows related columns
- X column, since it seems an index column
- user column.
We keep the columns of actual sensor data.

```{r message=FALSE}
na_cols <- colSums(is.na(trainingRAW))!=0
remove_cols <- grepl("timestamp|window|user|X",names(trainingRAW))
training <- trainingRAW[,!(na_cols|remove_cols)]
```
## Data Processing and Training
We create data partition in the training set to perform the training in a part and validate the model in the other part. We chose 0.4 to get around 7500 observations.
More observations will increase the processing time of the model.
Now it takes less than an hour in a core i5 with 8GB RAM

```{r cache=TRUE, message=FALSE}
library(caret)
inTrain <- createDataPartition(training$classe,p=0.4,list=FALSE)
subtraining <- training[inTrain,]
crossvalid <- training[-inTrain,]
modelFit <- train(classe ~.,data=subtraining,method="rf")
```

## Model Validation
We validate the model predicting the values of the classe variable of the observations not used to compute the model. Then we compare with the real values of the observations to calculate the Accuracy and other error measurements:
```{r}
library(caret)
confusionMatrix(crossvalid$classe,predict(modelFit,crossvalid))
```
## Predict testing values
We can know calculate the prediction of the classe variable for the testing dataset.
```{r}
result <-predict(modelFit,testingRAW)
```
